Mozilla Data Science Project - 2019

### 2019-12-17 poc/EDA/EDA.Rmd
updating the EDA report
finalizing EDA and comparative analysis

# Exploratory Data Analysis (EDA)

This R markdown covers the Exploratory Data Analysis (EDA) on the training dataset of the PRD.

## Setting-up
```{r message=FALSE, warning=FALSE}
## Loading the needed libraries

library(kableExtra)      # help you build common complex tables and manipulate table styles
library(tidyverse)       # for general data wrangling (includes readr and dplyr)
library(ggplot2)         # to draw statistical plots 
library(plotly)          # to construct interactive 3d plots
library(DataExplorer)    # automated data exploration
library(corrplot)        # to plot nice correlation matrix
library(caret)           # includes several functions to pre-process
library(scales)          # to determining breaks and labels for axes and legends
library(skimr)
library(funModeling) 
library(Hmisc)
library(grid)
library(hrbrthemes)
library(tidyr)
library(viridis)
library(ggpubr)
library(ggthemes)
library(GGally)
library(nortest)
```

```{r message=FALSE, warning=FALSE}
## Loading the training dataset

@@

@@

@@

## Step 1 : Data Inspection

@@

```{r echo=FALSE, fig.height=10, fig.width=10}
plot_intro(df_train_f, ggtheme = theme_minimal())
```

@@

```{r echo=FALSE, fig.height=10, fig.width=10}
plot_intro(df_validate_f, ggtheme = theme_minimal())
```

* Most of the dataset is composed of continuous variables
* No NAs values (were handled in preprocessing)

## Data Structure

@@

Let's use `glimpse` function to display a vertical preview of the validation dataset. So we can easily preview data type and sample data.

@@

@@

## Step 2 : Beta vs Release

```{r, fig.width = 7.5}
df_release <- df_train_f[which(df_train_f$label == 'release'), ]
df_beta <- df_train_f[which(df_train_f$label == 'beta'), ]

f <- freq(df_train_f$label)
```

```{r}
summary(df_release)
```

```{r}
summary(df_beta)
```

```{r, fig.width = 7.5}
df_v_release <- df_validate_f[which(df_validate_f$label == 'release'), ]
df_v_beta <- df_validate_f[which(df_validate_f$label == 'beta'), ]

f <- freq(df_validate_f$label)
```

```{r}
summary(df_v_release)
```

```{r}
summary(df_v_beta)
```

## Step 3 - Analyzing Discrete Variables

```{r, fig.height=15, fig.width=15, message=FALSE, warning=FALSE}
## Frequency distribution release dataframe
plot_bar(df_release, ggtheme = theme_minimal(base_size = 15))
```

```{r, fig.height=15, fig.width=15, message=FALSE, warning=FALSE}
## Frequency distribution beta dataframe
plot_bar(df_beta, ggtheme = theme_minimal(base_size = 15))
```

```{r, fig.height=15, fig.width=15, message=FALSE, warning=FALSE}
## Frequency distribution release dataframe
plot_bar(df_v_release, ggtheme = theme_minimal(base_size = 15))
```

```{r, fig.height=15, fig.width=15, message=FALSE, warning=FALSE}
## Frequency distribution beta dataframe
plot_bar(df_v_beta, ggtheme = theme_minimal(base_size = 15))
```

## Step 4 - Analyzing Continuos Variables

```{r, fig.width = 15, fig.height = 15}
## View histogram of release dataset
plot_histogram(df_release, ggtheme = theme_minimal(base_size = 15))
```

```{r, fig.width = 15, fig.height = 15}
## View histogram of beta dataset
plot_histogram(df_beta, ggtheme = theme_minimal(base_size = 15))
```

```{r, fig.width = 15, fig.height = 15}
## View histogram of release dataset
plot_histogram(df_release, ggtheme = theme_minimal(base_size = 15))
```

```{r, fig.width = 15, fig.height = 15}
## View histogram of beta dataset
plot_histogram(df_beta, ggtheme = theme_minimal(base_size = 15))
```

@@

### Ploting Density Curves

```{r include=FALSE}
require(cowplot)
```

@@

# User Engagement Continuous Metrics

@@

@@

## Comparing Two Continuous Distributions

The QQ plot can be used to compare two continuous distributions.

@@

  test <- ks.test(x, y)$statistic
  pvalue <- ks.test(x, y)$p.value

  test <- paste("KS Test = ", round(test, 3))
  pvalue <- paste("P-value = ", round(pvalue, 3))

  qqplot(x, y, main=i, xlim=rg, ylim=rg, xlab = "Beta", ylab = "Release", pch = 1)
  # mtext(test, side=3)
  text(min(x), max(x), paste(pvalue, "\n", test), adj=c(0,1))
  abline(0,1, col="red")  
}
```

```{r message=FALSE, warning=FALSE, out.width=c('50%', '50%')}
for (i in user_eng) {
  x <- df_beta_ue[,i]
  y <- df_release_ue[,i]

  print(i)
  # print(ad.test(x))
  # print(ad.test(y))
  print(ks.test(x, y))
}
```

The QQ plot can be used to compare two continuous distributions.

@@

  test <- ks.test(x, y)$statistic
  pvalue <- ks.test(x, y)$p.value

  test <- paste("KS Test = ", round(test, 3))
  pvalue <- paste("P-value = ", round(pvalue, 3))

  qqplot(x, y, main=i, xlim=rg, ylim=rg, xlab = "Beta", ylab = "Release", pch = 1)
  # mtext(test, side=3)
  text(min(x), max(x), paste(pvalue, "\n", test), adj=c(0,1))
  abline(0,1, col="red")  
}
```

```{r message=FALSE, warning=FALSE, out.width=c('50%', '50%')}
for (i in user_eng) {
  x <- df_beta_v_ue[,i]
  y <- df_release_v_ue[,i]

  print(i)
  # print(ad.test(x))
  # print(ad.test(y))
  print(ks.test(x, y))
}
```

@@

* Analyzing only the plots, in general, the distributions are very similar to each other
* Mainly the metrics related to **active hours** and **number of active days**, **search count**, **number of pages**, **daily unique domains** and **daily number of sessions started**
* Comparing only the training and validation plots, we can notice that the variables that presented a more different behavior were: `num_active_days`, `uri_count_max` and `num_addons`
* When I performed the KS (kolmogorov-Smirnov) test, the null hypothesis that both samples come from the same distribution was rejected for all the comparisons. *Am I doing something wrong?*

# User Engagement Discrete Metrics

@@

## Comparing Two Discrete Distributions

@@

  rel_beta <- table(x)/nrow(df_beta_ue_dis) #divide the frequency counts by the total
  beta_bar <- barplot(rel_beta,
        main = "Beta", #Give your chart a title
        ylim=c(0,1), border=F, col = "navy",
        xlab = i, #Label the x axis
        ylab = "Relative Frequency" #Label the y axis
  ) 
  # Add the text 
  text(beta_bar, rel_beta+0.025, paste(round(rel_beta*100), "%", sep="") ,cex=1)

  rel_release <- table(y)/nrow(df_release_ue_dis) #divide the frequency counts by the total
  release_bar <- barplot(rel_release,
        main = "Release", #Give your chart a title
        ylim=c(0,1), border=F, col = "navy",
        xlab = i, #Label the x axis
        ylab = "Relative Frequency" #Label the y axis
  )

@@

  rel_beta <- table(x)/nrow(df_beta_v_ue_dis) #divide the frequency counts by the total
  beta_bar <- barplot(rel_beta,
        main = "Beta", #Give your chart a title
        ylim=c(0,1), border=F, col = "navy",
        xlab = i, #Label the x axis
        ylab = "Relative Frequency" #Label the y axis
  ) 
  # Add the text 
  text(beta_bar, rel_beta+0.025, paste(round(rel_beta*100), "%", sep="") ,cex=1)

  rel_release <- table(y)/nrow(df_release_v_ue_dis) #divide the frequency counts by the total
  release_bar <- barplot(rel_release,
        main = "Release", #Give your chart a title
        ylim=c(0,1), border=F, col = "navy",
        xlab = i, #Label the x axis
        ylab = "Relative Frequency" #Label the y axis
  )

@@

@@

* Analyzing only the plots, in general, the distributions are very similar to each other

>>>>

# Introduction

The goal of this project is to utilize [statistical matching](https://en.wikipedia.org/wiki/Matching_(statistics)) techniques to search for a subset of the prerelease user population that is most representative of Release.

In this first report, we will perform an exploratory analysis of the data, focusing on investigating the differences between Beta and Release users.

# Loading the data

```{r message=FALSE, warning=FALSE, include=FALSE}
## Loading the needed libraries

library(kableExtra)      # help you build common complex tables and manipulate table styles
library(tidyverse)       # for general data wrangling (includes readr and dplyr)
library(ggplot2)         # to draw statistical plots 
library(plotly)          # to construct interactive 3d plots
library(DataExplorer)    # automated data exploration
library(corrplot)        # to plot nice correlation matrix
library(caret)           # includes several functions to pre-process
library(scales)          # to determining breaks and labels for axes and legends
library(skimr)
library(funModeling) 
library(Hmisc)
library(grid)
library(hrbrthemes)
library(tidyr)
library(viridis)
library(ggpubr)
library(ggthemes)
library(GGally)
library(nortest)
library(emdist)
```

```{r echo=TRUE, message=FALSE, warning=FALSE}
## Loading the training dataset

@@

* App Version 67

@@

* App Version 68

@@

# Data Inspection

@@

@@

------

## Observations

```{r echo=FALSE, fig.height=10, fig.width=10}
par(mfrow = c(2, 2))  ## Set up a 2 x 2 plotting space

plot_intro(df_train_f, ggtheme = theme_minimal(), title = "Training")
plot_intro(df_validate_f, ggtheme = theme_minimal(), title = "Validation")
```

* In both datasets, most variables are continuous
* No NAs value (were handled in preprocessing)

# Data Structure

@@

Let's use `glimpse` function to display a vertical preview of the validation dataset.

@@

------

## Observations

@@

# Beta vs Release

```{r}
## Training
df_release <- df_train_f[which(df_train_f$label == 'release'), ]
df_beta <- df_train_f[which(df_train_f$label == 'beta'), ]

## Validation
df_v_release <- df_validate_f[which(df_validate_f$label == 'release'), ]
df_v_beta <- df_validate_f[which(df_validate_f$label == 'beta'), ]
```

* Composition:

```{r}
f <- freq(df_train_f$label)
```

* Composition:

```{r}
f <- freq(df_validate_f$label)
```

# Analyzing Discrete Variables

```{r, fig.height=15, fig.width=15, message=FALSE, warning=FALSE}
## Frequency distribution release dataframe
plot_bar(df_release, ggtheme = theme_minimal(base_size = 15), title = 'Release')
```

```{r, fig.height=15, fig.width=15, message=FALSE, warning=FALSE}
## Frequency distribution beta dataframe
plot_bar(df_beta, ggtheme = theme_minimal(base_size = 15), title = 'Beta')
```

```{r, fig.height=15, fig.width=15, message=FALSE, warning=FALSE}
## Frequency distribution release dataframe
plot_bar(df_v_release, ggtheme = theme_minimal(base_size = 15), title = 'Release')
```

```{r, fig.height=15, fig.width=15, message=FALSE, warning=FALSE}
## Frequency distribution beta dataframe
plot_bar(df_v_beta, ggtheme = theme_minimal(base_size = 15), title = 'Beta')
```

# Analyzing Continuos Variables

```{r, fig.width = 15, fig.height = 15}
## View histogram of release dataset
plot_histogram(df_release, ggtheme = theme_minimal(base_size = 15), title = 'Release')
```

```{r, fig.width = 15, fig.height = 15}
## View histogram of beta dataset
plot_histogram(df_beta, ggtheme = theme_minimal(base_size = 15), title = 'Beta')
```

```{r, fig.width = 15, fig.height = 15}
## View histogram of release dataset
plot_histogram(df_release, ggtheme = theme_minimal(base_size = 15), title = 'Release')
```

```{r, fig.width = 15, fig.height = 15}
## View histogram of beta dataset
plot_histogram(df_beta, ggtheme = theme_minimal(base_size = 15), title = 'Beta')
```

------

## Observations

@@

## Ploting Density Curves

```{r message=FALSE, warning=FALSE, include=FALSE}
require(cowplot)
```

@@

# User Engagement Metrics

## Continuous Variables

@@

#### Beta-Release Difference

@@

### QQ-Plots

#### Training

@@

  test <- ks.test(x, y)$statistic
  # pvalue <- ks.test(x, y)$p.value

  test <- paste("KS Test = ", round(test, 3))
  # pvalue <- paste("P-value = ", round(pvalue, 3))

  qqplot(x, y, main=i, xlim=rg, ylim=rg, xlab = "Beta", ylab = "Release", pch = 1)
  # mtext(test, side=3)
  text(min(x), max(x), test, adj=c(0,1))
  abline(0,1, col="red", lty=2)  
}
```

#### Validation

@@

  test <- ks.test(x, y)$statistic
  # pvalue <- ks.test(x, y)$p.value

  test <- paste("KS Test = ", round(test, 3))
  # pvalue <- paste("P-value = ", round(pvalue, 3))

  qqplot(x, y, main=i, xlim=rg, ylim=rg, xlab = "Beta", ylab = "Release", pch = 1)
  # mtext(test, side=3)
  text(min(x), max(x), test, adj=c(0,1))
  abline(0,1, col="red", lty=2)  
}
```

------

@@

```{r message=FALSE, warning=FALSE, include=FALSE}
tbl_ks <- data.frame(
training = c(0.071,0.045,0.044,0.047,0.054,0.099,0.073,0.014,0.019,0.026,0.055,0.054,0.601,0.029,0.033,0.094,0.080,0.035,0.032,0.119,0.116,0.129,0.018,0.042,0.208,0.075,0.064,0.046,0.025), 
validation = c(0.166,0.058,0.077,0.071,0.095,0.076,0.040,0.045,0.059,0.042,0.082,0.082,0.292,0.040,0.054,0.075,0.051,0.036,0.053,0.098,0.117,0.130,0.017,0.035,0.235,0.109,0.087,0.062,0.035)
)

row.names(tbl_ks) <- c('num_active_days','active_hours','active_hours_max','uri_count','uri_count_max','session_length','session_length_max','search_count','search_count_max','num_bookmarks','num_pages','num_pages_max','num_addons','daily_unique_domains','daily_unique_domains_max','daily_max_tabs','daily_max_tabs_max','daily_tabs_opened','daily_tabs_opened_max','daily_num_sessions_started','daily_num_sessions_started_max','startup_ms','install_year','profile_age','timezone_offset','memory_mb','cpu_cores','cpu_speed_mhz','cpu_l2_cache_kb')
```

```{r}
kable(tbl_ks) %>%
add_header_above(c("KS Distance" = 3)) %>%
kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F) %>%
row_spec(c(2,18,21:26), bold = T, color = "white", background = "#c3f584")
```

* Analyzing at the plots of each dataset separately, in general, Beta-Release distributions are very similar to each other
* For instance, metrics related to **active hours** and **number of active days**, **search count**, **number of pages**, **daily unique domains** and **daily number of sessions started**
* Comparing only the training and validation plots, we can observe large differences between the variables. Only a few showed similar distances (marked in green)

## Discrete Variables

This section will focus only on user engagement discrete metrics. So, we are going to analyze the following metrics:

@@

@@

  rel_beta <- table(x)/nrow(df_beta_ue_dis) #divide the frequency counts by the total
  beta_bar <- barplot(rel_beta,
        main = "Beta", #Give your chart a title
        ylim=c(0,1), border=F, col = "#ffd271",
        xlab = i, #Label the x axis
        ylab = "Relative Frequency" #Label the y axis
  ) 
  # Add the text 
  text(beta_bar, rel_beta+0.025, paste(round(rel_beta*100), "%", sep="") ,cex=1)

  rel_release <- table(y)/nrow(df_release_ue_dis) #divide the frequency counts by the total
  release_bar <- barplot(rel_release,
        main = "Release", #Give your chart a title
        ylim=c(0,1), border=F, col = "#f65c78",
        xlab = i, #Label the x axis
        ylab = "Relative Frequency" #Label the y axis
  )

@@

  rel_beta <- table(x)/nrow(df_beta_v_ue_dis) #divide the frequency counts by the total
  beta_bar <- barplot(rel_beta,
        main = "Beta", #Give your chart a title
        ylim=c(0,1), border=F, col = "#ffd271",
        xlab = i, #Label the x axis
        ylab = "Relative Frequency" #Label the y axis
  ) 
  # Add the text 
  text(beta_bar, rel_beta+0.025, paste(round(rel_beta*100), "%", sep="") ,cex=1)

  rel_release <- table(y)/nrow(df_release_v_ue_dis) #divide the frequency counts by the total
  release_bar <- barplot(rel_release,
        main = "Release", #Give your chart a title
        ylim=c(0,1), border=F, col = "#f65c78",
        xlab = i, #Label the x axis
        ylab = "Relative Frequency" #Label the y axis
  )

@@

------

@@

* In general, the distributions are very similar to each other for both versions
